Automated Topic Naming

### 2011-08-23 ese/ese.tex
added some text on pg numbers and timeline

%\noindent \textbf{Timeline of key events} -- 
We analyzed each project's developer mailing list for external validation. 
%As discussed earlier, the top row in each figure shows key events for each project. 
We use \textit{labelled topic extraction} to pick out the underlying NFR activity behind these events. 
For example, both projects show a high number of NFRs recognized at the first period of analysis. 
This is due to our window choice: we deliberately targeted our
analysis to when both MySQL 3.23
and MaxDB 7.500 were first announced. For MaxDB, version 7.5.00  was released in December of 2003. 
We know that release 7.5.00.23 saw the development of PHP interfaces, possibly accounting for the simultaneous increase in the \emph{portability}
NFR at the same time.
The gap in MaxDB (Figure \ref{fig:maxdb-timeline}) is due to a shift in development focus (from February 2005 to June 2005) to MaxDB 7.6, which is
released in June 2005.

%For MySQL (Figure \ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data. 
The release of MySQL we study  (Figure \ref{fig:mysql-timeline}) was the first to be licenced under the GPL. 
Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labelled with \emph{functionality} and
\emph{maintainability}. 
After this point, this version enters the maintenance phase of its life-cycle. 
In May 2001, there is an increase in the number of topics labelled with \emph{portability}. 
This might be related to release 3.23.38, which focused on Windows compatibility. 
Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release
of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model). 
After this point, efforts shift to the newer releases (4.0, 4.1,
5.0). 
\todo{add PGSQL disucssion here. Schema additions? Lots of functionality?}

@@

\subsection{Do NFR Frequencies Change over Time?}
%\noindent\textbf{1. Do NFR frequencies change over time?} --- 
%\subsubsection{Do label frequencies change over time?}
\todo{add more on PGSQL}
In both projects the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
%As discussed in the previous section, 
Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.

\subsection{Do Projects Differ in their Relative  Topic Interest?}
%\subsubsection{Do  projects differ in their relative  topic interest?}
%\noindent\textbf{2. Do  projects differ in their relative  topic interest?} ---
Yes. MySQL 3.23 had proportionally more
topics labelled \emph{functionality}, while MaxDB had proportionally more
\emph{efficiency} related topics. MaxDB was a very mature release ``donated'' to the open-source community, 
whereas MySQL was in its relative infancy, and	
security problems were more common (security is a component of \emph{functionality} in the ISO9126 model). 
In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers
how often portability arises as a concern.
%On one hand, this analysis is unsurprising; on the other hand, it provides confirmation that our approach identifies actual patterns.
%\subsection{Summary}

%\todo[inline]{link the actual data, which is interesting, to somethign
 % actionable for a developer or manager}

\subsection{Do different developers work on different NFRs?}
\todo{figure this out!}

@@

%XXX can just remove this
%\subsection{Effectiveness}
\subsection{Annotation Observations}
We found many topics that were not actually non-functional requirements (NFRs) but were often related to them. 
For instance, concurrency was mentioned often in the commit logs and
was related to \emph{correctness} and \emph{reliability}, possibly because concurrent code is prone to bugs such as race conditions. %are 
% often  was troublesome. 
\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration
management. 
these kinds of changes are slightly related to \emph{maintainability}. 
A non-functional change that was not quality-related was licensing and
copyright; many changes were simply to do with updating copyrights or
ensuring copyright or license headers were applied to files. In these
cases we assigned the \emph{None} label to the topic.

>>>>

%\noindent \textbf{Timeline of key events} -- 
We analyzed each project's developer mailing list for external validation. 
%As discussed earlier, the top row in each figure shows key events for each project. 
We use \textit{labelled topic extraction} to pick out the underlying NFR activity behind these events. 
For example, both MaxDB and MySQL show a high number of NFRs recognized at the first period of analysis. 
This is due to our window choice: we deliberately targeted our
analysis to when both MySQL 3.23
and MaxDB 7.500 were first announced. For MaxDB, version 7.5.00  was released in December of 2003. 
We know that release 7.5.00.23 saw the development of PHP interfaces, possibly accounting for the simultaneous increase in the \emph{portability}
NFR at the same time.
The gap in MaxDB (Figure \ref{fig:maxdb-timeline}) is due to a shift in development focus (from February 2005 to June 2005) to MaxDB 7.6, which is
released in June 2005.

%For MySQL (Figure \ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data. 
The release of MySQL we study  (Figure \ref{fig:mysql-timeline}) was the first to be licenced under the GPL. 
Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labelled with \emph{functionality} and
\emph{maintainability}. 
After this point, this version enters the maintenance phase of its life-cycle. 
In May 2001, there is an increase in the number of topics labelled with \emph{portability}. 
This might be related to release 3.23.38, which focused on Windows compatibility. 
Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release
of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model). 
After this point, efforts shift to the newer releases (4.0, 4.1,
5.0).

By contrast, the PostgreSQL timeline  (Figure \ref{fig:pgsql-timeline}) is extracted from a central trunk and is not version-specific. Therefore development 
tends to be focused on releases 7.3, 7.4 and 7.5/8.0 alpha. For example, a priority for the 8.0 candidate was a Windows-native port of the source code, which seems to 
correlate with the \emph{portability} NFR increasing in frequency in mid-2004. Before the 7.4 release, \emph{usability,functionality} and \emph{portability} all 
increase in frequency, possibly reflecting the interest in adding features and documentation for the release.

@@

\subsection{Do NFR Frequencies Change over Time?}
%\noindent\textbf{1. Do NFR frequencies change over time?} --- 
%\subsubsection{Do label frequencies change over time?}
%\todo{add more on PGSQL}
In both MaxDB and MySQL the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
%As discussed in the previous section, 
Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.

In PostgreSQL, by comparison, the frequencies seem to by cyclic, since we are not studying a maintenance-phase for the product, but rather ongoing
feature addition and usability improvements.

\subsection{Do Projects Differ in their Relative  Topic Interest?}
Yes. MySQL 3.23 had proportionally more
topics labelled \emph{functionality}, while MaxDB had proportionally more
\emph{efficiency} related topics. MaxDB was a very mature release ``donated'' to the open-source community, 
whereas MySQL was in its relative infancy, and	
security problems were more common (security is a component of \emph{functionality} in the ISO9126 model). 
In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the lifetime of the projects. It may surprise developers
how often portability arises as a concern.
\todo{add the pgsql data}

\subsection{Do different developers work on different NFRs?}
We did a t-test to assess whether there was statistically significant differences in the types of NFRs developers chose to work on, for the PostgreSQL project only. 
\todo{add more, figure}

@@

\subsection{Annotation Observations}
We found many topics that were not actually non-functional requirements (NFRs) but were often related to them. 
For instance, concurrency was mentioned often in the commit logs and
was related to \emph{correctness} and \emph{reliability}, possibly because concurrent code is prone to bugs such as race conditions. %are 
% often  was troublesome. 
\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration
management. 
these kinds of changes are slightly related to \emph{maintainability}. 
A non-functional change that was not quality-related was licensing and
copyright; many changes were simply to do with updating copyrights or
ensuring copyright or license headers were applied to files. In these
cases we assigned the \emph{None} label to the topic.
